---
title: "Prototype for AMC Laos shiny app"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About

This document contains development code for the AMC Laos shiny app.

# Preparation

```{r packages}
library(DT)
library(glue)
library(highcharter)
library(janitor)
library(leaflet)
library(readxl)
library(rgdal)
library(tidyverse)
library(validate)
```


# Data

## Data Import

```{r data}
amc_dta <- read_excel(path = "./www/data/AMC_data_2021-03-10.xlsx")
amc_dta <- clean_names(amc_dta)

# sheet ACT: removed some content in columns outside the table.
amc_dic <- list(
  act = read_excel(path = "./www/data/AMC_Dictionary_2021-03-10.xlsx", sheet = "ACT") %>% clean_names(),
  act3 = read_excel(path = "./www/data/AMC_Dictionary_2021-03-10.xlsx", sheet = "ACT3 & ATC name") %>% clean_names(),
  atc5 = read_excel(path = "./www/data/AMC_Dictionary_2021-03-10.xlsx", sheet = "ACT3 & 5, substance and AWaRe") %>% clean_names()
)

coords <- read.csv("./www/data/hospital_geo_coordinates.csv", stringsAsFactors = FALSE)

shp_lao_provinces <- readOGR("./www/data/shapefiles/provinces.shp")
```


## Data Check

```{r data-check}
# Columns should be predicatbles
all(names(amc_dta) == c("provinces", "data_collecting_year", "atc5", "route", "dad", 
                        "act3", "a_wa_re", "substance", "act_3_name", "dad_percent"))


valid_atc5  <- unique(amc_dic$atc5$atc5)
valid_act3  <- unique(amc_dic$act3$code)
valid_name <- unique(amc_dic$act3$name)
valid_substance  <- unique(amc_dic$atc5$substance)

rules <- validator(
  is.character(provinces),
  is.numeric(data_collecting_year),
  is.character(atc5),
  is.character(route),
  is.numeric(dad),
  is.character(act3),
  is.character(a_wa_re),
  is.character(substance),
  is.character(act_3_name),
  is.numeric(dad_percent),
  provinces %in% c("Luang Natha", "Xiengkhuang", "Salavan"),
  atc5 %in% valid_atc5,
  route %i% c("O", "P", "R"),
  act3 %in% valid_act3,
  a_wa_re %in% c("Access", "Watch"),
  substance %in% valid_substance,
  act_3_name %in% valid_name,
  in_range(dad_percent, min = 0, max = 100)
)

confront(amc_dta, rules)
summary(confront(amc_dta, rules))

# Issue with three substances ----
unique(amc_dta$substance)[! unique(amc_dta$substance) %in% valid_substance]
# "Choramphenicol" instead of "Chloramphenicol"
# "Amoxicillin and enzyme inhibitor"  "Sulfamethoxazole and trimethoprim" 

amc_dta %>% 
  group_by(provinces) %>%
  summarise(add_pct = sum(dad_percent))
```


## Data Wrangling

```{r}
amc_dta <- amc_dta %>%
  mutate(hospital = case_when(
    provinces == "Luang Natha" ~ "Luang Namtha Provincial Hospital",
    provinces == "Xiengkhuang" ~ "Xiengkhuang Provincial Hospital",
    provinces == "Salavan" ~ "Salavan Provincial Hospital"
  ))
```


# Visualisations

Map of hospitals:

```{r}
dta <- amc_dta %>%
  group_by(hospital) %>%
  summarise(nb_year = n_distinct(data_collecting_year)) %>%
  mutate(label = glue("{hospital}: {nb_year} year(s) of AMC data"))

dta <- left_join(dta, coords, by = c("hospital" = "hospcd"))

leaflet(data = dta) %>% 
  addTiles() %>%
  addPolygons(data = shp_lao_provinces, fill = FALSE, dashArray = 1, weight = 0, color='black') %>%
  addMarkers(lng = ~lon, lat = ~lat, label = ~label, labelOptions = labelOptions(noHide = T))
```

Number of consumed antimicrobials by hospital by year:

```{r}
dta <- amc_dta %>%
  group_by(data_collecting_year, hospital) %>%
  summarise(consum = round(sum(dad), 1), .groups = "drop")

hchart(dta, type = "column", hcaes(x = data_collecting_year, y = consum, group = hospital)) %>%
  hc_yAxis(title = "Total DAD") %>% 
  hc_xAxis(title = "Year, Hospital") %>%
  hc_tooltip(headerFormat = "", pointFormat = "{point.consum} DAD<br> {point.hospital} {point.data_collecting_year}")
```


Antimicrobial consumptions based on AWaRE group:

```{r}
dta <- amc_dta %>%
  group_by(a_wa_re) %>%
  summarise(tot = sum(dad))


highchart() %>%
  hc_add_series(dta, "pie", hcaes(name = a_wa_re, y = tot),
                dataLabels = list(enabled = TRUE, style = list(fontSize = "13px"),
                                  format = '{point.name} <br>{point.percentage:.0f} %')) %>%
  hc_title(text = "AWaRe Group") %>%
  hc_tooltip(headerFormat = "", pointFormat = "{point.tot:.1f} DAD")
```


Route of administration:

```{r}
dta <- amc_dta %>%
  group_by(route) %>%
  summarise(tot = sum(dad))


highchart() %>%
  hc_add_series(dta, "pie", hcaes(name = route, y = tot),
                dataLabels = list(enabled = TRUE, style = list(fontSize = "13px"),
                                  format = '{point.name} <br>{point.percentage:.2f} %')) %>%
  hc_title(text = "Route of administration") %>%
  hc_tooltip(headerFormat = "", pointFormat = "{point.tot:.1f} DAD")
```

Common consumed antimicrobial group:

```{r}
dta <- amc_dta %>%
  group_by(act_3_name) %>%
  summarise(consum = round(sum(dad), 1), .groups = "drop") %>%
  arrange(desc(consum))

hchart(dta, type = "column", hcaes(x = act_3_name, y = consum)) %>%
  hc_yAxis(title = "Total DAD") %>% 
  hc_xAxis(title = "ACT3 - Antibiotic Group") %>%
  hc_tooltip(headerFormat = "", pointFormat = "{point.consum} DAD")
```

Common consumed antimicrobial agents:

```{r}
dta <- amc_dta %>%
  group_by(substance) %>%
  summarise(consum = round(sum(dad), 1), .groups = "drop") %>%
  arrange(desc(consum))

hchart(dta, type = "column", hcaes(x = substance, y = consum)) %>%
  hc_yAxis(title = "Total DAD") %>% 
  hc_xAxis(title = "Substance") %>%
  hc_tooltip(headerFormat = "", pointFormat = "{point.consum} DAD")
```




List of antimicrobial consumptions:

```{r}
dta <- amc_dta %>%
  group_by(act_3_name, substance, route, a_wa_re) %>%
  summarise(consum = round(sum(dad), 1), .groups = "drop") %>%
  arrange(desc(consum))

datatable(dta)
```



